---
title: "Dropout Data (Part II)"
author: "Cristal Quiñones"
date: "7/8/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

A continuation of "Dropout Data" to keep it from getting too large. 

### Sunday work: making bar graphs 

The first step is really to figure out how many dropouts there were TOTAL for each of our academic years. Then, we can figure out how to see how many of those dropouts are male and how many are female. So I guess technically the first step is to make a new data set with all this information, but to do this, I will first do the data frames with counties and total drop out for each year. 
```{r}
# First step is to make similar data frames for all years
# I've already done 2011 and 2012, so all that is left is 2013, 2014, 2015 and 2016

# Remove scientific notaion 
options(scipen=999)
#2013 
thirteendata$County2013 <- thirteendata$CDS_CODE
thirteendata$County2013 <- substr(thirteendata$County2013, 1,2)

thirteendata$SumDo <- thirteendata$DTOT + thirteendata$D7 + thirteendata$D8

ListOfCounties2013 <- as.numeric(unique(thirteendata$County2013)) 
CountiesTotals2013 <- matrix(NA, nrow=length(ListOfCounties2013), ncol=2)
for(i in 1:length(ListOfCounties2013)){ 
  rowi <- which(thirteendata$County2013 == ListOfCounties2013[i])  
  CountiesTotals2013[i,2] = sum(thirteendata[rowi,22]) 
  CountiesTotals2013[i,1] = ListOfCounties2013[i] 
}
CountiesTotals2013 # printing our new data frame 
```

```{r}
# 2014
library(stringr)
options(scipen=999)

fourteendata$County2014 <- fourteendata$CDS_CODE
fourteendata$County2014 <- substr(fourteendata$County2014, 1,2)

fourteendata$SumDo2014 <- fourteendata$DTOT + fourteendata$D7 + fourteendata$D8

ListOfCounties2014 <- as.numeric(unique(fourteendata$County2014))
CountiesTotals2014 <- matrix(NA, nrow=length(ListOfCounties2014), ncol=2)
for(j in 1:length(ListOfCounties2014)){
  rowj <- which(fourteendata$County2014 == ListOfCounties2014[j])
  CountiesTotals2014[j,2] = sum(fourteendata[rowj,23])
  CountiesTotals2014[j,1] = ListOfCounties2014[j]
}
CountiesTotals2014
```

```{r}
fifteendata$County2015 <- fifteendata$CDS_CODE
fifteendata$County2015 <- substr(fifteendata$County2015, 1,2)

fifteendata$SumDo2015 <- fifteendata$DTOT + fifteendata$D7 + fifteendata$D8

ListOfCounties2015 <- as.numeric(unique(fifteendata$County2015))
CountiesTotals2015 <- matrix(NA, nrow=length(ListOfCounties2015), ncol=2)
for(j in 1:length(ListOfCounties2015)){
  rowj <- which(fifteendata$County2015 == ListOfCounties2015[j])
  CountiesTotals2015[j,2] = sum(fifteendata[rowj,23])
  CountiesTotals2015[j,1] = ListOfCounties2015[j]
}
CountiesTotals2015
```

```{r}
fifteendata$County2015 <- fifteendata$CDS_CODE
fifteendata$County2015 <- substr(fifteendata$County2015, 1,2)

fifteendata$SumDo2015 <- fifteendata$DTOT + fifteendata$D7 + fifteendata$D8

ListOfCounties2015 <- as.numeric(unique(fifteendata$County2015))
CountiesTotals2015 <- matrix(NA, nrow=length(ListOfCounties2015), ncol=2)
for(j in 1:length(ListOfCounties2015)){
  rowj <- which(fifteendata$County2015 == ListOfCounties2015[j])
  CountiesTotals2015[j,2] = sum(fifteendata[rowj,23])
  CountiesTotals2015[j,1] = ListOfCounties2015[j]
}
CountiesTotals2015
```

```{r}
sixteendata$County2016 <- sixteendata$CDS_CODE
sixteendata$County2016 <- substr(sixteendata$County2016, 1,2)

sixteendata$SumDo2016 <- sixteendata$DTOT + sixteendata$D7 + sixteendata$D8

ListOfCounties2016 <- as.numeric(unique(sixteendata$County2016))
CountiesTotals2016 <- matrix(NA, nrow=length(ListOfCounties2016), ncol=2)
for(j in 1:length(ListOfCounties2016)){
  rowj <- which(sixteendata$County2016 == ListOfCounties2016[j])
  CountiesTotals2016[j,2] = sum(sixteendata[rowj,22])
  CountiesTotals2016[j,1] = ListOfCounties2016[j]
}
CountiesTotals2016
```


Now we have the six data sets that contain the county and their respective dropout numbers. The next step is to have a total for each year. This would require taking the sum of the second column for each one of the data frames we've just made. 

```{r}
YearDORate <- matrix(NA, nrow=1, ncol=2)
for(i in 1:58){
  YearDORate[1,2] = sum(CountiesTotals2011[i,2])
  YearDORate[1,1] = "Total"
}
```

```{r}
years <- c(2011:2016)
total2011 <- sum(CountiesTotals2011[,2])
total2012 <- sum(CountiesTotals2012[,2])
total2013 <- sum(CountiesTotals2013[,2])
total2014 <- sum(CountiesTotals2014[,2])
total2015 <- sum(CountiesTotals2015[,2])
total2016 <- sum(CountiesTotals2016[,2])

yearlytotals <- c(total2011, total2012, total2013, total2014, total2015, total2016)

yearlytotalsDF <- data.frame(years, yearlytotals)
names(yearlytotalsDF) <- c("Years", "Totals")
```

Okay so now (in a really roundabout way), we have a data matrix with the years and their totals. Now, we make a bargraph with it. And then we have to figure out how many of them are female and how many are male. 

```{r}
# Simple bargraph 
library(ggplot2)

dobar <- ggplot(yearlytotalsDF, aes(x=Years, y=Totals)) + geom_bar(stat="identity") + labs(title="Yearly Dropout Rates") + scale_x_discrete(limits = c(2011, 2012,2013,2014,2015, 2016))
```

```{r}
options(scipen=999)
library(ggplot2)
library(scales)
# Load data sets 
elevendata <- read.csv(file="dropout2011_12.csv", header = T)
twelvedata <- read.csv(file="dropouttwelve.csv", header = T)
thirteendata <- read.csv(file="dropout2013_14.csv", header = T)
fourteendata <- read.csv(file="dropout2014_15.csv", header = T)
fifteendata <- read.csv(file="dropout2015_16.csv", header = T)
sixteendata <- read.csv(file="dropout2016_17.csv", header = T)

elevendata <- elevendata[,-c(1)]
#twelvedata <- twelvedata[,-c(1)]
#thirteendata <- thirteendata[,-c(1)]
fourteendata <- fourteendata[,-c(1)]
fifteendata <- fifteendata[,-c(1)]
#sixteendata <- sixteendata[,-c(1)]

elevendata$SumDO <- elevendata$D7 + elevendata$D8 + elevendata$DTOT
twelvedata$SumDO <- twelvedata$D7 + twelvedata$D8 + twelvedata$DTOT
thirteendata$SumDO <- thirteendata$D7 + thirteendata$D8 + thirteendata$DTOT
fourteendata$SumDO <- fourteendata$D7 + fourteendata$D8 + fourteendata$DTOT
fifteendata$SumDO <- fifteendata$D7 + fifteendata$D8 + fifteendata$DTOT
sixteendata$SumDO <- sixteendata$D7 + sixteendata$D8 + sixteendata$DTOT
# Print old column names 

# Same column numbers 
new2011 <- elevendata[, c(1:3, 20,21)]
new2012 <- twelvedata[, c(1:3, 20,21)]
new2013 <- thirteendata[,c(1:3,20,21)]
new2014 <- fourteendata[,c(1:3,20,21)]
new2015 <- fifteendata[,c(1:3,20,21)]
new2016 <- sixteendata[,c(1:3,20,21)]

# Same column names 
colnames(new2011) <- c("CDS_CODE","ETHNIC","GENDER","YEAR","SUM")
colnames(new2012) <- c("CDS_CODE","ETHNIC","GENDER","YEAR","SUM")
colnames(new2013) <- c("CDS_CODE","ETHNIC","GENDER","YEAR","SUM")
colnames(new2014) <- c("CDS_CODE","ETHNIC","GENDER","YEAR","SUM")
colnames(new2015) <- c("CDS_CODE","ETHNIC","GENDER","YEAR","SUM")
colnames(new2016) <- c("CDS_CODE","ETHNIC","GENDER","YEAR","SUM")

# Merging all data sets 
combined <- rbind(new2011, new2012, new2013, new2014, new2015, new2016)
# Renaming years - change to 2011-12 when you have time 
combined$YEAR <- as.character(combined$YEAR)
combined$YEAR[combined$YEAR == "1112"] <- "2011-12"
combined$YEAR[combined$YEAR == "1213"] <- "2012-13"
combined$YEAR[combined$YEAR == "1314"] <- "2013-14"
combined$YEAR[combined$YEAR == "1415"] <- "2014-15"
combined$YEAR[combined$YEAR == "1516"] <- "2015-16"
combined$YEAR[combined$YEAR == "1617"] <- "2016-17"

combined$GENDER <- as.character(combined$GENDER)
combined$GENDER[combined$GENDER == "F"] <- "Female"
combined$GENDER[combined$GENDER == "M"] <- "Male"
# Bar plot with genders 
# taking the sums so we get the total number of female students dropping out and total nubmber f male students dropping out
newcombinedgen <- combined[,-c(1,2)]
geninfo <- aggregate(.~YEAR+GENDER, data=newcombinedgen, sum)

GenPlot <- ggplot(geninfo, aes(YEAR,SUM)) + geom_bar(aes(fill=GENDER), stat="identity") + labs(title="Yearly Dropout Counts in California K-12 Schools", subtitle = "By Gender", x="Years", y="Total", fill="Gender") + theme(plot.title = element_text(hjust=0.5), plot.subtitle = element_text(hjust = 0.5)) + scale_y_continuous(limits=c(0,90000), breaks = seq(0,90000,10000), labels = comma)

# Relabeling so that ethnicities are represented by name and not by number
combined$ETHNIC <- as.character(combined$ETHNIC)
combined$ETHNIC[combined$ETHNIC == 5] <- "Hispanic/Latinx"
combined$ETHNIC[combined$ETHNIC == 6] <- "Black"
combined$ETHNIC[combined$ETHNIC == 7] <- "White"
combined$ETHNIC[combined$ETHNIC == 2] <- "Asian"
combined$ETHNIC[combined$ETHNIC == 0] <- "Other"
combined$ETHNIC[combined$ETHNIC == 1] <- "Other"
combined$ETHNIC[combined$ETHNIC == 3] <- "Other"
combined$ETHNIC[combined$ETHNIC == 4] <- "Other"
combined$ETHNIC[combined$ETHNIC == 9] <- "Other"
# Bar plot with ethnicities
newcombinedeth <- combined[,-c(1,3)]
ethinfo <- aggregate(.~YEAR+ETHNIC, data=newcombinedeth,sum)


EthPlot <- ggplot(ethinfo, aes(YEAR,SUM)) + geom_bar(aes(fill=ETHNIC), stat="identity") + labs(title="Yearly Dropout Counts in California K-12 Schools", subtitle = "By Race/Ethnicity", x="Years", y="Total", fill="Race/Ethnicity") + theme(plot.title = element_text(hjust=0.5), plot.subtitle = element_text(hjust = 0.5)) + scale_y_continuous(limits=c(0, 90000), breaks=seq(0,90000,10000), labels = comma)


write.csv(ethinfo, file="/Users/owner/ethinfo.csv") # in theory this means you should only have to make one like to read these two files and the plots 
write.csv(geninfo, file="/Users/owner/geninfo.csv")
```

To add angle to x axis, include 
axis.text.x = element_text(angle = 45)
in the theme 
```{r}
#Trying to change it to percent form 
EthPercentONE <- ggplot(ethinfo, aes(YEAR,(ethinfo$SUM)/sum(ethinfo$SUM))) + geom_bar(aes(fill=ETHNIC), stat="identity") 
``` 

New Data sets.

We downloaded new data: graduate outcome data, and two different povery data sets. 
```{r}
#loading graduation data into R
grad1314 <- read.csv(file="grads1314.csv", header=TRUE)
grad1415 <- read.csv(file="grads1415.csv", header=TRUE)
grad1516 <- read.csv(file="grads1516.csv", header=TRUE)
grad1617 <- read.csv(file="grads1617.csv", header=TRUE)

#loading unduplicated student poverty data // free reduced lunch data 
frpm1314 <- read.csv(file="frpm1314.csv", header=TRUE)
frpm1415 <- read.csv(file="frpm1415.csv", header=TRUE)
frpm1516 <- read.csv(file="frpm1516.csv", header=TRUE)
frpm1617 <- read.csv(file="frpm1617.csv", header=TRUE)

#loading other poverty data 

```

