---
title: "Education (MSRI)"
author: "Cristal Quiñones"
date: "7/3/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---
Summaries for Dropouts in 2011-12 
```{r}
elevendata <- read.csv(file="dropout2011_12.csv", header=TRUE)

elevendata <- elevendata[,-c(1)] #removing the first column which is just an index 

# Col 1 - Schools 
length(unique(elevendata$CDS_CODE)) # Finding how many unique elements are in the column CDS_CODE - returns 5382
length((elevendata$CDS_CODE)) # checking the length of the original column

#now we do the same for all 20 columns  

# Col 2 - ethnicities 
length(unique(elevendata$ETHNIC)) # returns 9 

# Col 3 - gender 
length(unique(elevendata$GENDER)) # 2 

sum(elevendata$E7) #465108
sum(elevendata$E8) #468275
sum(elevendata$E9) # 501258
sum(elevendata$E10) # 495009
sum(elevendata$E11) # 487466 
sum(elevendata$E12) # 495945
sum(elevendata$EUS) #5096 
sum(elevendata$ETOT) # 1984774
sum(elevendata$D7) #3367
sum(elevendata$D8) #3023
sum(elevendata$D9) # 8394
sum(elevendata$D10) # 10874
sum(elevendata$D11) # 15237 
sum(elevendata$D12) # 44589
sum(elevendata$DUS) # 319
sum(elevendata$DTOT) #79413


```

Dropout 2012 - 2013 # does not exist? 
```{r}
twelvedata <- read.csv(file="dropout2012_13.csv", header=TRUE)

length(unique(twelvedata$CDS_CODE)) #936

# Col 2 - ethnicities 
length(unique(twelvedata$ETHNIC)) # 9 
# Col 3 - gender 
length(unique(twelvedata$GENDER)) # 2 

sum(twelvedata$E7) #470494
sum(twelvedata$E8) #466132
sum(twelvedata$E9) # 497455
sum(twelvedata$E10) # 486498
sum(twelvedata$E11) # 481531 
sum(twelvedata$E12) # 499275
sum(twelvedata$EUS) #5271 
sum(twelvedata$ETOT) # 1970030
sum(twelvedata$D7) #3397
sum(twelvedata$D8) #3062
sum(twelvedata$D9) # 9362
sum(twelvedata$D10) # 10237
sum(twelvedata$D11) # 14993 
sum(twelvedata$D12) # 42606
sum(twelvedata$DUS) # 318
sum(twelvedata$DTOT) #77516
```

Dropout 2013 - 2014 also does not exist 
```{r}
thirteendata <- read.csv(file="dropout2013_14.csv", header=T)
length(unique(thirteendata$CDS_CODE))
length(unique(thirteendata$ETHNIC))

sum(thirteendata$E7) # 465592
sum(thirteendata$E8) # 472381
sum(thirteendata$E9) # 491493
sum(thirteendata$E10) # 484993
sum(thirteendata$E11) # 477425
sum(thirteendata$E12) # 498403
sum(thirteendata$EUS) # 5603
sum(thirteendata$ETOT) # 1957917
sum(thirteendata$D7) # 1671
sum(thirteendata$D8) # 1282 
sum(thirteendata$D9) # 5917 
sum(thirteendata$D10) # 6485 
sum(thirteendata$D11) #10710 
sum(thirteendata$D12) # 38292
sum(thirteendata$DUS) # 196
sum(thirteendata$DTOT) # 61600 

```

Dropout 2014 - 2015 
```{r}
fourteendata <- read.csv(file="dropout2014_15.csv", header=T)
length(unique(fourteendata$CDS_CODE)) # 5432
length(unique(fourteendata$ETHNIC)) # 9
# gender = 2 
sum(fourteendata$E7) # 464286
sum(fourteendata$E8) # 467038
sum(fourteendata$E9) # 495004
sum(fourteendata$E10) # 480753
sum(fourteendata$E11) # 477097
sum(fourteendata$E12) # 496901
sum(fourteendata$EUS) # 5600
sum(fourteendata$ETOT) # 1955355
sum(fourteendata$D7) # 1659
sum(fourteendata$D8) # 1284 
sum(fourteendata$D9) # 6165 
sum(fourteendata$D10) # 5720 
sum(fourteendata$D11) #8705 
sum(fourteendata$D12) # 32978
sum(fourteendata$DUS) # 236
sum(fourteendata$DTOT) # 53804 
```

Dropout 2015 - 2016 
```{r}
fifteendata <- read.csv(file="dropout2015_16.csv", header=TRUE)
length(unique(fifteendata$CDS_CODE)) # 5476
length(unique(fifteendata$ETHNIC)) # 9 
# gender is 2 

sum(fifteendata$E7) # 470753
sum(fifteendata$E8) # 465322
sum(fifteendata$E9) # 487202
sum(fifteendata$E10) # 488004
sum(fifteendata$E11) # 472968
sum(fifteendata$E12) # 492835
sum(fifteendata$EUS) # 6316
sum(fifteendata$ETOT) # 1947325
sum(fifteendata$D7) # 1506
sum(fifteendata$D8) # 1155 
sum(fifteendata$D9) # 6952 
sum(fifteendata$D10) # 5345 
sum(fifteendata$D11) #8057 
sum(fifteendata$D12) # 29906
sum(fifteendata$DUS) # 286
sum(fifteendata$DTOT) # 50546 
```

Dropout 2016-2017 
```{r}
sixteendata <- read.csv(file="dropout2016_17.csv", header=TRUE)
length(unique(sixteendata$CDS_CODE)) # 952
length(unique(sixteendata$ETHNIC)) # 9

sum(sixteendata$E7) # 473574
sum(sixteendata$E8) # 472340
sum(sixteendata$E9) # 487547
sum(sixteendata$E10) # 486085
sum(sixteendata$E11) # 481521
sum(sixteendata$E12) # 484169
sum(sixteendata$EUS) # 6079
sum(sixteendata$ETOT) # 1945401
sum(sixteendata$D7) # 1453
sum(sixteendata$D8) # 1138 
sum(sixteendata$D9) # 7698 
sum(sixteendata$D10) # 5743 
sum(sixteendata$D11) # 7873 
sum(sixteendata$D12) # 25641
sum(sixteendata$DUS) # 294
sum(sixteendata$DTOT) # 47249 
```

```{r}
# Now we plot 
library(ggplot2)
plot(twelvedata$E7, twelvedata$D7)
plot(twelvedata$ETHNIC, twelvedata$DTOT)
qplot(twelvedata$ETHNIC, twelvedata$DTOT)

p <- ggplot(data=elevendata, aes(x=ETHNIC, y=DTOT)) +
  geom_bar(stat="")
p

t <- ggplot(data=fifteendata, aes(x=ETHNIC, y=DTOT)) + geom_bar(stat="identity")
t
```

We are going to attempt to make a new data frame. We want to make it so that each school name gets one row, not 18, and the total sum is added. We know for one, eah school has 18 observations dedicated to it. There are 8 variables describing dropouts: D7, D8, D9, D10, D11, D12, DUS, DTOT. Since DTOT is the sum of a subset of the others, we will not include it in our calculations. Instead, we will be taking the 18 rows and 7 columns. 
```{r}
# we will only do the first two data sets 
# first i want to see if there is any code that shows you exactly where each school shows up. 
# according to the pdf about the data, each school may have 1 or more records (rows) but no more than 16 
# the first thing I've done to move in the right direction is remove the scientific notation so I can see the full code and more easily distinguish each school 
# we'll start small and then move into doing what we actually want. for example, let's first find out all the dropouts from the first school (code: 1100170109835)
totS1 <- sum(elevendata$D7[c(1:17)]) + sum(elevendata$D8[c(1:17)]) + sum(elevendata$D9[c(1:17)]) + sum(elevendata$D10[c(1:17)]) + sum(elevendata$D11[c(1:17)]) + sum(elevendata$D12[c(1:17)]) +sum(elevendata$DUS[c(1:17)])

# just realized that instead of adding every D column, I can just add DTOT to D7 and D8 since DTOT is defined as all droupouts exclusind years 7 and 8 
totS2 <- sum(elevendata$D7[c(1:17)]) + sum(elevendata$D8[c(1:17)]) + sum(elevendata$DTOT[c(1:17)])
totS2

# Now I know the simple command to add specifc rows in specific columns but there are a couple issues # (1) how do I make this not so tedius
# (2) how do I know which rows to add 

# we are going to do it by district 
library(stringr)

#elevendata$District <- elevendata$CDS_CODE
#elevendata$District <- as.character(elevendata$District)
#elevendata$District <- substr(elevendata$District, 3,7)

elevendata$county <- elevendata$CDS_CODE
elevendata$county <- substr(elevendata$county, 1,2)

elevendata$SumDo <- elevendata$DTOT + elevendata$D7 + elevendata$D8

# Now we have a column with the sum of all dropouts for that entry, one step closer to what we want 
# The goal is to  have a data frame with all 58 districts and the total number of dropouts they had 
# Let's find which rows hold information for what district 

# Summing the rows of total dropout for district 11 
row11 <- which(elevendata$county == 11)
sum(elevendata[row11,23])

#now we make a more general pieec of code 
counties <- as.numeric(unique(elevendata$county))
CountiesTotals2011 <- matrix(NA, nrow=length(counties), ncol=2)
for(i in 1:length(counties)){
  rowi <- which(elevendata$county == counties[i])
  CountiesTotals2011[i,2] = sum(elevendata[rowi,23])
  CountiesTotals2011[i,1] = counties[i]
}
CountiesTotals2011
#This was the original attempt which was not as good but a good starting point 
#dist1 <- 0
#for(i in 1:nrow(elevendata)){
 # if(elevendata$county[i] == "11"){
  #  dist1 <- dist1 + as.numeric(elevendata$SumDo[i])
  #} else {
  #NULL
  #}
#}

```

Now we can attempt to do the same with a different data set. This time with data from years 2012 - 2013 

```{r}
# We want to add the columns we need: SumDo and Counties

# Adding the column of counties 
twelvedata$County <- twelvedata$CDS_CODE
twelvedata$County <- substr(twelvedata$County, 1,2)

# Add column SumDo which is the total number of dropouts for that specific population in that school in the county we are looking at (ex: the total number of black boys that dropped at any year in school a in district 11)
twelvedata$SumDo <- twelvedata$DTOT + twelvedata$D7 + twelvedata$D8

# A for Loop to make a data frame containing every school county and the total number of dropouts they had in the academic year 2012 - 2013 

ListOfCounties2012 <- as.numeric(unique(twelvedata$County)) # A list of the unique county codes 
CountiesTotals2012 <- matrix(NA, nrow=length(ListOfCounties2012), ncol=2) # an empty matrix to be populated with the county codes and their respective dropout rates 
for(i in 1:length(ListOfCounties2012)){ 
  rowi <- which(twelvedata$County == ListOfCounties2012[i]) # finding the rows with the same county code 
  CountiesTotals2012[i,2] = sum(twelvedata[rowi,22]) # summing the total number of dropouts and populating our previously empty dataframe 
  CountiesTotals2012[i,1] = ListOfCounties2012[i] # labeling the dropout rates by their county 
}
CountiesTotals2012 # printing our new data frame 
```

### Sunday work: making bar graphs 

The first step is really to figure out how many dropouts there were TOTAL for each of our academic years. Then, we can figure out how to see how many of those dropouts are male and how many are female. So I guess technically the first step is to make a new data set with all this information, but to do this, I will first do the data frames with counties and total drop out for each year. 
```{r}
# First step is to make similar data frames for all years
# I've already done 2011 and 2012, so all that is left is 2013, 2014, and 2015 


```
