---
title: "Clean Code"
author: "Cristal Quiñones"
date: "7/19/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---
So, I have been using "as.integer" becuase I did not know not to, and now I have to redo my models. This is the code to do so. 
The things we will do here:
1. Model (One) Graduates with Dropouts, Suspensions, Expulsions, English Learner, Free/Reduced Lunch, and Foster
2. Model (Two) Graduates with every variables except graduate demographics 
3. Check collinearity for Model 1
4. Check collinearity for Model 2 (o)
5. Model One with standardized and normalized data 
6. Model Two with standardized and normalized data 
7. Model One broken down by gender with standardized and normalized data 
8. Model One broken down by ethnicity with standardized and normalized data 
9. For each one of the models, use stargazed to export a latex table with the coefficient, p-value, and standard error 

```{r}
library(dplyr)
library(stargazer)
library(readxl)
library(ggplot2)
library(GGally)
library(MASS)
library(corpcor)
library(mctest)
library(ppcor)
library(BBmisc)

sourceDF <- read.csv(file="Sus_Ex_Do_en_cp_ucCsu_DEMO.csv", header=T)
sourceDF <- sourceDF[,-c(1,38)] # remove index and county 
sourceDF$do_total <- sourceDF$do_female + sourceDF$do_male
write.csv(sourceDF, file="/Users/owner/sus_ex_en_cp_uccsu_DEMO.csv")

sourceDF <- na.omit(sourceDF)


#--------------------------Model One--------------------------------------
modelone <- lm(ucCsu_total~do_total+TotalSuspensions+Expulsion_Total+EnglishLearner+FreeReduced+Foster, data = sourceDF)

stargazer(modelone, dep.var.labels = c("UC/CSU Eligible Grduates"), covariate.labels = c("Dropout","Suspension","Expulsion","English Learner","Free or Reduced Lunch","Foster Youth"), report = ('vcps'), out="modelone.tex") # printing data table for latex 

#-------------------------Model Two---------------------------------------
modeltwoDF <- sourceDF[,-c(1, 12:18)] # data frame without dist_year and UC/CSU demographic variables
modeltwo <- lm(ucCsu_total~., data = modeltwoDF)
stargazer(modeltwo, dep.var.labels = c("UC/CSU Eligible Graduates"), covariate.labels = c("Enrolled (Asian)","Enrolled (Black)","Enrolled (Hispanic)","Enrolled (White)","Enrolled (Other)","Enrolled (Female)","Enrolled (Male)","English Learner","Free or Reduced Lunch","Foster Youth", "Enrolled","Suspension (Female)","Suspension (Male)","Suspension","Suspension (Asian)","Suspension (Black)","Suspension (Hispanic)","Suspension (White)","Suspension (Other)","Expulsion (Female)","Expulsion (Male)","Expulsion","Expulsion (Asian)","Expulsion (Black)","Expulsion (Hispanic)","Expulsion (White)", "Expulsion (Other)","Dropout (Asian)","Dropout (Black)","Dropout (Hispanic)","Dropout (White)", "Dropout (Other)","Dropout (Female)","Dropout (Male)","Dropout"), report = ('vc*p*s'), out="modeltwo.tex")

#-----------------------Model Two without Enrollment----------------------------
modeltwoDFNoEn <- modeltwoDF[,-c(1:7,11)] # data frame without district yr, UCCSU, or enroll
modeltwoNoEn <- lm(ucCsu_total~., data = modeltwoDFNoEn)
stargazer(modeltwoNoEn, dep.var.labels = c("UC/CSU Eligible Graduates"), covariate.labels = c("English Learner","Free or Reduced Lunch","Foster Youth","Suspension (Female)","Suspension (Male)","Suspension","Suspension (Asian)","Suspension (Black)","Suspension (Hispanic)","Suspension (White)","Suspension (Other)","Expulsion (Female)","Expulsion (Male)","Expulsion","Expulsion (Asian)","Expulsion (Black)","Expulsion (Hispanic)","Expulsion (White)", "Expulsion (Other)","Dropout (Asian)","Dropout (Black)","Dropout (Hispanic)","Dropout (White)", "Dropout (Other)","Dropout (Female)","Dropout (Male)","Dropout"), report = ('vc*p*s'), out="modeltwoNoEn.tex")

#-----------------------Check collinearity: Model One---------------------
X <- sourceDF[,c(44,23,31,9,10,11)] # the columns we used for our linear model 
ggpairs(X) # shows that EnglishLearner and FRLunch seem colinear 
cor2pcor(cov(X)) # correlation matrix shows pair-wise corr among explanatory var not high, except for english learners-Free Reduced Lunch
omcdiag(X, sourceDF$ucCsu_total) # the value of standardized determinant is .01 (small) indicating a strong degree of multicollinearity.Chi-square test statistic is 335.6651 which is significant which all indicates multicollinearity in the model 
imcdiag(X,sourceDF$ucCsu_total) # English Learners, Free/Reduced Lunch have pretty high F-Statistics, the strongest being Free/Reduced Lunch followed by english learner
pcor(X, method = "pearson") # high partial correlation between: Free Reduced Lunch & English Learners

#----------------------Model One addressing Collinearity (no FRLunch)-----
modeloneFRL <- lm(ucCsu_total~do_total+TotalSuspensions+Expulsion_Total+EnglishLearner+Foster, data = sourceDF)
stargazer(modelone, dep.var.labels = c("UC/CSU Eligible Grduates"), covariate.labels = c("Dropout","Suspension","Expulsion","English Learner","Foster Youth"), report = ('vc*p*s'), out="modelone.tex")

#----------------------Model One addressing Collinearity (no EL)----------
modeloneEL <- lm(ucCsu_total~do_total+TotalSuspensions+Expulsion_Total+FreeReduced+Foster, data = sourceDF)
stargazer(modelone, dep.var.labels = c("UC/CSU Eligible Grduates"), covariate.labels = c("Dropout","Suspension","Expulsion","Free or Reduced Lunch","Foster Youth"), report = ('vc*p*s'), out="modeloneEL.tex")

#-----------------------Check collinearity: Model Two---------------------
# not yet....too many variables

#-----------------------Relitavizing sourceDF Data-------------------------
# We're dividing every clumn by the corresponding enrollment column 
fr <- sourceDF

fr$EnglishLearnerrel <- fr$EnglishLearner/fr$enroll_total 
fr$FreeReducedrel <- fr$FreeReduced/fr$enroll_total
fr$Fosterrel <- fr$Foster/fr$enroll_total
fr$TotalSuspensionsrel <- fr$TotalSuspensions/fr$enroll_total
fr$do_totalrel <- fr$do_total/fr$enroll_total
fr$ucCsu_totalrel <- fr$ucCsu_total/fr$enroll_total
fr$Expulsion_Totalrel <- fr$Expulsion_Total/fr$enroll_total

fr$FemaleSuspensionsrel <- fr$FemaleSuspensions/fr$enroll_female
fr$do_femalerel <- fr$do_female/fr$enroll_female
fr$ucCsu_femalerel <- fr$ucCsu_female/fr$enroll_female
fr$Expulsion_Femalerel <- fr$Expulsion_Female/fr$enroll_female

fr$MaleSuspensionsrel <- fr$MaleSuspensions/fr$enroll_male
fr$do_malerel <- fr$do_male/fr$enroll_male
fr$ucCsu_malerel <- fr$ucCsu_male/fr$enroll_male
fr$Expulsion_Malerel <- fr$Expulsion_Male/fr$enroll_male

fr$AsianSuspensionrel <- fr$AsianSuspensions/fr$enroll_asian
fr$do_asianrel <- fr$do_asian/fr$enroll_asian
fr$ucCsu_asianrel <- fr$ucCsu_asian/fr$enroll_asian
fr$Expulsion_Asianrel <- fr$Expulsion_Asian/fr$enroll_asian

fr$BlackSuspensionsrel <- fr$BlackSuspensions/fr$enroll_black
fr$do_blackrel <- fr$do_black/fr$enroll_black
fr$ucCsu_blackrel <- fr$ucCsu_black/fr$enroll_black
fr$Expulsion_Blackrel <- fr$Expulsion_Black/fr$enroll_black

fr$HispanicSuspensionsrel <- fr$HispanicSuspensions/fr$enroll_hisp
fr$do_hisprel <- fr$do_hisp/fr$enroll_hisp
fr$ucCsu_hisprel <- fr$ucCsu_hisp/fr$enroll_hisp
fr$Expulsion_Hispanicrel <- fr$Expulsion_Hispanic/fr$enroll_hisp

fr$WhiteSuspensionsrel <- fr$WhiteSuspensions/fr$enroll_white
fr$do_whiterel <- fr$do_white/fr$enroll_white
fr$ucCsu_whiterel <- fr$ucCsu_white/fr$enroll_white
fr$Expulsion_Whiterel <- fr$Expulsion_White/fr$enroll_white

fr$OtherSuspensionsrel <- fr$OtherSuspensions/fr$enroll_total
fr$do_otherrel <- fr$do_other/fr$enroll_total
fr$ucCsu_otherrel <- fr$ucCsu_other/fr$enroll_total
fr$Expulsion_Otherrel <- fr$Expulsion_Other/fr$enroll_total

RELsourceDF <- fr[,c(1,45:79)] # to use instead of the regular standardized and normalized data 
write.csv(RELsourceDF, file="/Users/owner/relativeCompleteMerge.csv")

#------------------Standardizing and Normalizing sourceDF Data------------
dist_year <-as.character(sourceDF$dist_year) # we will append it after manipulating the variable values 
justvalues <- sourceDF[,-c(1)] #identical df sans dist_year 
zjustvalues <- scale(justvalues, center = TRUE, scale = TRUE) # scaled data https://stat.ethz.ch/R-manual/R-devel/library/base/html/scale.html 
nzjustvalues <- normalize(zjustvalues, method = 'range', range = c(0,1)) # norm data https://www.rdocumentation.org/packages/BBmisc/versions/1.10/topics/normalize

test1 <- as.data.frame(nzjustvalues) # converting the standardized and normalized just values into a data frame 
nzsourceDF <- cbind(dist_year,test1) # data frame with all the values and dist_year (as "factor")
write.csv(nzsourceDF, file="/Users/owner/nzsourceDF.csv")

#--------------------Model One (Standardized and Normalized)-------------
modeloneNZ <- lm(ucCsu_total~do_total+TotalSuspensions+Expulsion_Total+EnglishLearner+FreeReduced+Foster, data = nzsourceDF)
stargazer(modeloneNZ, type="latex", dep.var.labels = c("UC/CSU Eligible Graduates"), covariate.labels = c("Dropouts","Suspensions","Expulsions","English Learners","Free or Reduced Lunch","Foster Youth"), report = ('vc*p*s'), out = "modeloneNZ.tex")

#---------------------Model One (Relative Data)--------------------------
modeloneREL <- lm(ucCsu_totalrel~do_totalrel+TotalSuspensionsrel+Expulsion_Totalrel+EnglishLearnerrel+FreeReducedrel+Fosterrel, data = RELsourceDF)
stargazer(modeloneREL, type="latex", dep.var.labels = c("UC/CSU Eligible Graduates"), covariate.labels = c("Dropouts","Suspensions","Expulsions","English Learners","Free or Reduced Lunch","Foster Youth"), report = ('vc*p*s'), out = "modeloneREL.tex")

#--------------------Model Two (Standardized and Normalized)-------------
modeltwoDFNZ <- nzsourceDF[,-c(1,12:18)] # normalized data for model two 
modeltwoNZ <- lm(ucCsu_total~., data = modeltwoDFNZ)
modeltwoNZmb <- lm(modeltwoDFNZ$ucCsu_total~., data = modeltwoDFNZ[,c(-12)])
stargazer(modeltwoNZ, dep.var.labels = c("UC/CSU Eligible Graduates"), covariate.labels = c("Enrolled (Asian)","Enrolled (Black)","Enrolled (Hispanic)","Enrolled (White)","Enrolled (Other)","Enrolled (Female)","Enrolled (Male)","English Learner","Free or Reduced Lunch","Foster Youth", "Enrolled","Suspension (Female)","Suspension (Male)","Suspension","Suspension (Asian)","Suspension (Black)","Suspension (Hispanic)","Suspension (White)","Suspension (Other)","Expulsion (Female)","Expulsion (Male)","Expulsion","Expulsion (Asian)","Expulsion (Black)","Expulsion (Hispanic)","Expulsion (White)", "Expulsion (Other)","Dropout (Asian)","Dropout (Black)","Dropout (Hispanic)","Dropout (White)", "Dropout (Other)","Dropout (Female)","Dropout (Male)","Dropout"), report = ('vcps'), out="modeltwoNF.tex")

#----------------Model Two (Relative Data - No enrollment)-----------------------
modeltwoDFREL <- RELsourceDF[,-c(1,11,15,19,23,27,31,35)] # relative data without dist year, uccsu, or enroll(already not included in realative data frame)
modeltwoREL<- lm(ucCsu_totalrel~., data = modeltwoDFREL)
stargazer(modeltwoREL, dep.var.labels = c("UC/CSU Eligible Graduates"), report = ('vc*p*s'), out="modeltwoREL.tex")

#--------------------Model One by Gender (Stan & Norm)------------------
modeloneGEN <- lm(ucCsu_total~do_female+do_male+FemaleSuspensions+MaleSuspensions+Expulsion_Female+Expulsion_Male+EnglishLearner+FreeReduced+Foster, data=nzsourceDF)
stargazer(modeloneGEN, type = "latex", dep.var.labels = c("UC/CSU Eligible Graduates"), covariate.labels = c("Dropout (Female)","Dropout (Male)","Suspension (Female)","Suspension (Male)","Expulsion (Female)","Expulsion (Male)","English Learner","Free or Reduced Lunch","Foster Youth"), report = ('vc*p*s'), out="modeloneGENNZ.tex") # export model as latex table

#-------------------Model One by Gender (realtive data)----------------
modeloneGENREL <- lm(ucCsu_totalrel~do_femalerel+do_malerel+FemaleSuspensionsrel+MaleSuspensionsrel+Expulsion_Femalerel+Expulsion_Malerel+EnglishLearnerrel+FreeReducedrel+Fosterrel, data=RELsourceDF)
stargazer(modeloneGENREL, type = "latex", dep.var.labels = c("UC/CSU Eligible Graduates"), covariate.labels = c("Dropout (Female)","Dropout (Male)","Suspension (Female)","Suspension (Male)","Expulsion (Female)","Expulsion (Male)","English Learner","Free or Reduced Lunch","Foster Youth"), report = ('vc*p*s'), out="modeloneGENREL.tex") 

#-------------------Model One by Ethnicity (Stan & Norm)----------------
modeloneETH <- lm(ucCsu_total~do_asian+do_black+do_hisp+do_white+do_other+Expulsion_Asian+Expulsion_Black+Expulsion_Hispanic+Expulsion_White+Expulsion_Other+AsianSuspensions+BlackSuspensions+HispanicSuspensions+WhiteSuspensions+OtherSuspensions+EnglishLearner+FreeReduced+Foster,data=nzsourceDF)
stargazer(modeloneETH, type="latex",dep.var.labels = c("UC/CSU Eligible Graduates"), covariate.labels = c("Dropout (Asian)","Dropout (Black)","Dropout (Hispanic)","Dropout (White)","Dropout (Other)", "Expulsion (Asian)","Expulsion (Black)","Expulsion (Hispanic)","Expulsion (White)","Expulsion (Other)", "Suspension (Asian)","Suspension (Black)","Suspension (Hispanic)","Suspension (White)","Suspension (Other)","English Learner","Free or Reduced Lunch","Foster Youth"), report = ('vc*p*s'), out = "modeloneethnicNZ.tex")

#-----------------Model One by Ethnicity (Realtive Data)---------------
modeloneETHREL <- lm(ucCsu_totalrel~do_asianrel + do_blackrel + do_hisprel + do_whiterel + do_otherrel + Expulsion_Asianrel + Expulsion_Blackrel + Expulsion_Hispanicrel + Expulsion_Whiterel + Expulsion_Otherrel + AsianSuspensionrel + BlackSuspensionsrel + HispanicSuspensionsrel + WhiteSuspensionsrel + OtherSuspensionsrel + EnglishLearnerrel + FreeReducedrel + Fosterrel ,data=RELsourceDF)

stargazer(modeloneETHREL, type="latex",dep.var.labels = c("UC/CSU Eligible Graduates"), covariate.labels = c("Dropout (Asian)","Dropout (Black)","Dropout (Hispanic)","Dropout (White)","Dropout (Other)", "Expulsion (Asian)","Expulsion (Black)","Expulsion (Hispanic)","Expulsion (White)","Expulsion (Other)", "Suspension (Asian)","Suspension (Black)","Suspension (Hispanic)","Suspension (White)","Suspension (Other)","English Learner","Free or Reduced Lunch","Foster Youth"), report = ('vc*p*s'), out = "modeloneethnicREL.tex")

#-----------------------Model One Fixed for Collinearity (Standardized)----------
modeloneELnz <- lm(ucCsu_total~do_total+TotalSuspensions+Expulsion_Total+FreeReduced+Foster, data = nzsourceDF)
stargazer(modelone, dep.var.labels = c("UC/CSU Eligible Grduates"), covariate.labels = c("Dropout","Suspension","Expulsion","Free or Reduced Lunch","Foster Youth"), report = ('vc*p*s'), out="modeloneELnz.tex")

```

```{r}
#-------------------------------Checking % graduated per district----------------
yearGradEnrollDF <- sourceDF[,c(1,40,44)]
yearGradEnrollDF$year <- substr(yearGradEnrollDF$dist_year, 7,13)
yearGradEnrollDF$dist <- as.numeric(substr(yearGradEnrollDF$dist_year, 1,5)) 
yearGradEnrollDF <- yearGradEnrollDF[,-c(1)]

GradAgg <- aggregate(ucCsu_total~dist, data = yearGradEnrollDF, sum)
EnrollAgg <- aggregate(enroll_tot~dist, data = yearGradEnrollDF, sum)

GradEnrollTot <- merge(GradAgg, EnrollAgg, by="dist")
GradEnrollTot <- transform(GradEnrollTot, p = (ucCsu_total / enroll_tot)) 
GradEnrollTot$p <- GradEnrollTot$p*100

# the only issue with this data frame is that the number of years per district varies but if it's a percent maybe it doesn't matter?

write.csv(GradEnrollTot, file="/Users/owner/percentgraduated.csv")

``` 


Now, we will do Decision trees. 
We will do this with relative data that does not contain enrollment information (model two) 
```{r}
library(tree)
library(randomForest)
library(rpart)
library(rpart.plot)

#tree1 <- tree(ucCsu_totalrel~., data = modeltwoDFREL)
#plot(tree1)
#text(tree1, cex = .3)
#summary(tree1)

rpart1 <- rpart(ucCsu_totalrel~., data = modeltwoDFREL, method = "anova")
printcp(rpart1)
plotcp(rpart1)

plot(rpart1, uniform = T, main = "Regression Tree")
text(rpart1, use.n = T, all = T, cex = 0.8)

pruned <- prune(rpart1, cp=0.04)
plot(pruned)
text(pruned, cex = .75, use.n = F, all = T, label=c('test'))
prp(pruned, faclen = 5)


plot(modeltwoDFREL$MaleSuspensionsrel, modeltwoDFREL$ucCsu_totalrel) # plot next to the tree - there is no linearity but the deicsion tree tels you there is a realtionship between male suspension and graduates that linear models do not tell you but you could explore further 

plot <- ggplot(modeltwoDFREL, aes(MaleSuspensionsrel, ucCsu_totalrel)) + geom_point() + labs(title = "Male Suspensions vs. UC/CSU Eligible Graduates", subtitle = "Using Enrollment Normalized Data", x = "Male Suspensions", y = "UC/CSU Eligible Graduates") + theme(plot.title = element_text(hjust=0.5), plot.subtitle = element_text(hjust = 0.5)) 

plot(plot)

```
